Depression and suicidality as costly signals of need

### 2021-03-31 Paper.rmd
Edited methods: - removed Pilot Studies section (MG draft; EH still in SI) - Improved wording on the design and vignette sections - minor edits and proofreading fixes elsewhere

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emohasize conflit and PI in the methods (at least in the pilot studies portion)
add abstract
add conclusion

@@

This study involved a vignette study with between subjects pre-test/post-test design with one factor (signal) and 5 levels (verbal request, crying, crying with behavioral changes, depression, and suicide attempt), with the outcome of interest being how believe and likelihood of helping changed in response to a given signal.

## Pilot Studies

The first pilot study used a vignette study to examine how the presence of conflict, private information, and emotional signaling impacted the degree participants believed the signaler to be in need and provide support using an MTurk sample of 1,447 Americans. Of the signals, depression resulted in the most pro-signaler responses (reword), separating itself from the less costly signals when conflict and private information was present and substantially outperforming the schizophrenic break control. However, the suicide attempt condition was responded to relatively poorly, contrary to predictions.

One potential explanation is that the initial scenario of a mother asking their sibling for help paying for their young daughter's heart surgery was not severe enough for a suicide attempt to be a functional signal, due to a lack of beleivability or due to its costs appearing manipulative. For this reason, we designed this study with more severe scenarios that were also associated with suicide attempts both ethnographically and clinically, including instances of sexual assasult, non-sexual assault, and thwarted marriage.

add that we ask now ask same question twice? (interpretabilty)

The second was posted on the samplesize subreddit with the goals of testing the beleivability of the new vignettes and how pro-signaler participants were at T1, with a power analysis with simulated data showing this would limit our power. In this small sample of 28, we saw no evidence that beleivabilty was a problem and no evidence that participants as a whole would be too pro-sister at T1.

@@

Four vignettes were used in this study, with participants in the US sample being randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a unversity basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and an imagined 15-year-old daughter after they lost their house in a fire. During this time, the participant's daughter becomes jealous of your niece before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette hard participants imagine that they found the best partner they have ever had after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. At one point, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is trying to arrange a dowry for their older daughter so she can marry the man she loves. After his family demands more money, the participant's family tries to find another man, who the older daughter claims to find unattractive. However, it is unclear how unattractive he is and any additional investment in the dowry will come at the expense of the younger daughter.

The full vignettes are available in the supplementary information.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than.

@@

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools, with there being some indication that it might outperform university samples in a few important areas. This is seen in that while both MTurk workers and university students engage in problematic behaviors that might lessen data quality to similar amounts [@NeckaEtAl2016], workers on MTurk have been shown to be more attentive than samples of university students [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from potential bot use and respondents faking their location to take surveys in a language they do not understand well, with the former concern being largely unfounded and the latter able to be corrected for with the tools used in this study [@KennedyEtAl2018]

@@

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2), 8) an attention check that did not filter out participants, 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

@@

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (BeliefT1) and the likelihood of them helping the signaler as requestedc (ActionT1). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (Divide T1). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. In every instance the position started at the left due to findings that people may be less likely to move the slider when it starts in the center (cite and word better). For the exact wording of the question and the ends of each continuum, see table 1.

@@

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (Action T3). In the US sample, this involved video evidence of the event in question occuring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the Action T3 question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment.

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable jump in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second, smaller jump from depression to a suicide attempt. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lenghty period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need (want to bring in MA here. Is that a bad idea?).

@@

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are encouraging. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception (cite), with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need (Hagen 2003), suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

>>>>

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emphasize conflict and PI in the methods (at least in the pilot studies portion)
add abstract
add conclusion

@@

This study involved a between-subjects pretest-posttest design to examine how different emotional signals would influence (1) the degree participants believed a fictional character to be in need and (2) the likelihood they would provide help within a vignette study.

@@

Four vignettes were used in this study in which both the presence of conflicts of interest and the existence of private information about one's need were emphasized. Due to the expectation that depression and suicide are likely too costly of signals for relatively minor instances of need, we designed this study to involve severe scenarios associated with suicide attempts both ethnographically and clinically (i.e, sexual assault (cite), non-sexual assault (cite), and thwarted marriage (cite)).

Participants in the US sample were randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a university basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and their 15-year-old daughter after they lost their house in a fire. During this time, the participant's daughter becomes jealous of her cousin before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette had participants imagine that they found a highly desirable partner after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. After a period of sustained conflict, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry the man she loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. However, it is unclear if the second man is as unattractive as the older daughter says or if she does not value her younger sister enough, since any additional investment in the dowry will come at the expense of the younger daughter.

The full vignettes are available in the SI.

@@

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] (check this. he might have a paper saying it's not associated with depression as people tend to think) and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than.

@@

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools, with there being some indication that it might outperform university samples in a few important areas. This is seen in that while both MTurk workers and university students engage in problematic behaviors that might lessen data quality to similar amounts [@NeckaEtAl2016], workers on MTurk have been shown to be more attentive than samples of university students [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from potential bot use and respondents faking their location to take surveys in a language they do not understand well, with the former concern being largely unfounded (at least for full on bots? do we mention human-assisted?)[@KennedyEtAl2018] and the risk of the latter able to be minimized through well designed attention checks (cite) .

@@

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2-T3), 8) an attention check that did not filter out participants (T4), 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

@@

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (T1Belief) and the likelihood of them helping the signaler as requested (T1Action). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (T1Divide). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. In every instance the position started at the left due to findings that people may be less likely to move the slider from central starting positions (cite and word better). For the exact wording of the question and the labels on the sliders, see table 1.

@@

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (T3Action). In the US sample, this involved video evidence of the event in question occurring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the T3Action question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment.

@@

Consistent with costly signaling theory (is this a common enough prediction to say this?), costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable jump in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second, smaller jump from depression to a suicide attempt. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lengthy period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need.

@@

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are encouraging (reword). In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception (cite), with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need (Hagen 2003), suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.
